```{r echo=FALSE}
rm(list=ls())
set.seed(1)
library(randomForest)
library(caret)
```

Practical Machine Learning Project
========================================================

## Introduction

Human Activity Recognition - HAR - has emerged as a key research area in the last years. There are many potential applications for HAR, like: elderly monitoring, life log systems for monitoring energy expenditure and for supporting weight-loss programs, and digital assistants for weight lifting exercises.

In this study six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E). Data were recorded with sensors in the participant's body and in the dumbbell.

We are going to use data generated by the sensors to predict the manner in which they did the exercise, this is the "classe" variable in the training set.


## Read data

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
setwd("C:/proyectos/cursos/practical machine learning/project")
data <- read.csv("pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))
dataTest <- read.csv("pml-testing.csv", na.strings = c("NA", "#DIV/0!", ""))
#data <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings = c("NA", #"#DIV/0!", ""))
#dataTest <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings = c("NA", #"#DIV/0!", ""))
```

## Preprocess data

A preview of the data set shows that there's a lot of columns with a majority of blank cells therefore we have delete this columns.


```{r}
isnaCount <- apply(data, 2, function(x){ sum(is.na(x))})
naColumns <- which(isnaCount > 0)
booleanNAColumns <- names(data) %in% names(naColumns)

data <- data[!booleanNAColumns]
dataTest <- dataTest[!booleanNAColumns]
```

Intentamos predecir que tipo de movimiento se está realizando en base a los datos aportados por los distintos sensores colocados en el cuerpo del sujeto, por tanto, variables como nombre del sujeto, fecha... no aportan ninguna información 

We are trying to predict the type of exercise that has been performed base on the data generated by the different sensors placed in the subject's body, therefore variables such as user_name or timestamp do not provide any information.


```{r}
data <- data[,8:60]
dataTest <- dataTest[,8:60]
```

## Cross validation

We are going to split the data in a training set (70% of the original set) and a test set.

```{r}
inTrain <- createDataPartition(y=data$classe, p=0.7, list=FALSE)
data <- data[inTrain,]
validationData <- data[-inTrain,]
```


## Fitting the model

Basándonos en el trabajo de "Velloso et al." escogemos random forest como el algoritmo apropiado para nuestro modelo. Al tratarse de un conjunto de datos bastante voluminoso escogemos un número de árboles reducido 

Based on the work of "Velloso et al." we use a random forest approach as the appropiate algorithm for our model. 
Becouse it is a large set of data we select a reduced number of trees (ntree=200) in order to increase 
performance.

```{r}
modFit <- randomForest(classe ~ ., data=data, importance=TRUE, ntree=200)
```

## Accuracy

To test the accuracy of the model we do an estimation using the test data and we compare the result with the real one using a table.

```{r }
p <- predict(modFit, validationData)
table(p, validationData$classe)
```

We can see in the table that there's no misclasified values, so we can stablish an accuracy of 100% on the test. data.