```{r echo=FALSE}
rm(list=ls())
set.seed(1)
library(randomForest)
library(caret)
```

Practical Machine Learning Project
========================================================

## Introduction


## Read data

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
data <- read.csv("pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))
dataTest <- read.csv("pml-testing.csv", na.strings = c("NA", "#DIV/0!", ""))
#data <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings = c("NA", #"#DIV/0!", ""))
#dataTest <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings = c("NA", #"#DIV/0!", ""))
```

## Preprocess data

Un vistazo previo al conjunto de datos nos muestra que existen muchas columnas con la mayoría de sus celdas en blanco, por tanto, eliminamos esas columnas.

A preview of the data set shows that there's a lot of columns with a majority of blank cells. So we delete this columns.


```{r}
isnaCount <- apply(data, 2, function(x){ sum(is.na(x))})
naColumns <- which(isnaCount > 0)
booleanNAColumns <- names(data) %in% names(naColumns)
data <- data[!booleanNAColumns]
dataTest <- dataTest[!booleanNAColumns]
```

Intentamos predecir que tipo de movimiento se está realizando en base a los datos aportados por los distintos sensores colocados en el cuerpo del sujeto, por tanto, variables como nombre del sujeto, fecha... no aportan ninguna información 

We are trying to predict the manner in which the exercise was done ... So variables like user_name or timestamp ...


```{r}
data <- data[,8:60]
dataTest <- dataTest[,8:60]
```

## Cross validation

We are going to use 70% of the data set to fit the model

```{r}
inTrain <- createDataPartition(y=data$classe, p=0.7, list=FALSE)
data <- data[inTrain,]
validationData <- data[-inTrain,]
```


## Fitting the model



```{r cache=TRUE}
modFit <- randomForest(classe ~ ., data=data, importance=TRUE, ntree=200)
```

## Accuracy

```{r }
p <- predict(modFit, validationData)
table(p, validationData$classe)
```